---
title: "Assignment 4: Matrix Algebra - Introduction to Computational Finance and Financial Econometrics"
author: "T Xu"
date: "Friday, September 26, 2014"
output: html_document
---
Questions 1 - 9 refer to the following matrices and vectors: 
$$
A= \left(\begin{array}\
1&4&7\\
2&4&8\\
6&1&3\\
\end{array}\right),
B= \left(\begin{array}\
4&4&0\\
5&9&1\\
2&2&5\\
\end{array}\right),
x= \left(\begin{array}\
1\\
2\\
3\\
\end{array}\right),
x= \left(\begin{array}\
5\\
2\\
7\\
\end{array}\right)
$$

Question 1
-----------------
(10) Compute the transpose of A.
```{r}
A <- matrix(c(1,4,7,2,4,8,6,1,3),ncol=3,nrow=3,byrow = T)
t(A)
```
Question 2
----------
(10) Compute the transpose of B.
```{r}
B <- matrix(c(4,4,0,5,9,1,2,2,5),ncol=3,nrow=3,byrow = T)
t(B)
```
Question 3
-------------
(5) Compute the transpose of x.
```{r}
x <- matrix(c(1,2,3))
t(x)
```
Question 4
---------------
(5) Compute the transpose of y.
```{r}
y <- matrix(c(5,2,7))
t(y)
```
Question 5
------------
(10) Compute A+B.
```{r}
A+B
```
Question 6
----------------
(10) Compute A−B.
```{r}
A-B
```

Question 7
---------
(10) Compute 2∗A
```{r}
2*A
```
Question 8
-----------
(10) Compute Ax.
```{r}
A %*% x
```

Question 9
-----------
(10) Compute y′Ax.
```{r}
t(y) %*% A %*% x
```

Question 10
------------
(10) Consider the system of equations: 
$$z1+z2=1,$$
$$2z1+4z2=2$$

Write the system using matrix notation as $Az=b$ and solve for z.
```{r}
A <- matrix(c(1,2,1,4),ncol=2)
b <- matrix(c(1,2))
solve(A,b) #calcualte z
```

Question 11
------------------------
Consider creating a portfolio of three assets denoted A, B and C. Assume the following information: 
$$
\mu= \left(\begin{array}\
0.01\\
0.04\\
0.02\\
\end{array}\right),
\Sigma= \left(\begin{array}\
0.10 & 0.30 & 0.10\\
0.30 & 0.15 & -0.20\\
0.10 & -0.20 & 0.08\\
\end{array}\right)
$$
(10) Compute the expected return for an equally weighted portfolio (i.e., xA=xB=xC=1/3).
```{r}
mu <- matrix(c(0.01,0.04,0.02))
sigma <- matrix(c(0.1,.3,.1,.3,.15,-.2,.1,-.2,0.08),ncol=3)
x <- matrix(c(1/3,1/3,1/3))
t(x) %*% mu
```


Question 12
-------------------
(10) Continuing from the previous question, what is the variance for an equally weighted portfolio?
```{r}
t(x) %*% sigma %*% x
```
